Supporting information - Bayesian regression facilitates quantitative modelling of cell metabolism

### 2024-02-12 paper/main.qmd
Review responses
Add callout blocks for new material

Maud uses Markov Chain Monte Carlo to perform posterior inference on a Bayesian
statistical model. This section introduces these topics and then describes
exactly what kind of Bayesian statistical model Maud uses.

@@

Maud employs adaptive Hamiltonian Monte Carlo to perform posterior inference
for a Bayesian statistical model where the process model is the kinetic model
described above. The next two subsections describe Maud's prior and measurement
models.

@@

@@

Lars Keld Nielsen: Supervision, writing manuscript, designing statistical model,
biological model, software architecture and user interface.   # References

@@

The authors declare no competing financial interest

>>>>

<!--:::{.callout-important title="Changes"}-->
Maud uses Markov Chain Monte Carlo to perform posterior inference on a Bayesian
statistical model. This section introduces these topics and then describes
exactly what kind of Bayesian statistical model Maud uses.

@@

Maud employs adaptive Hamiltonian Monte Carlo to perform posterior inference
for a Bayesian statistical model where the process model is the kinetic model
described above. The next two subsections describe Maud's prior and measurement
models.
<!--:::-->

@@

<!--:::{.callout-important title="Changes"}-->

@@

Lars Keld Nielsen: Supervision, writing manuscript, designing statistical model,
biological model, software architecture and user interface.

@@

The authors declare no competing financial interest.
<!--:::-->

# References

### 2024-02-12 paper/supporting_information.qmd
Add callout blocks for new material

While we have not yet observed this in practice, we expect that Maud is capable
of accurately sampling mildly multi-modal posterior distributions, i.e. those
for which regions of parameter space with high probability density are not
very sharply separated. This is because this ability depends primarily on the
underlying inference algorithm, and adaptive Hamiltonian Monte Carlo is known
to be able to sample many such posterior distributions. See, for example,
@mangiolaSccompRobustDifferential2023, which reports the successful use of
adaptive Hamiltonian Monte Carlo to sample posterior distributions involving
mixtures of Gaussian distributions, which are typically multi-modal.

@@

@zamora-silleroEfficientCharacterizationHighdimensional2011 introduces
the method HYPERSPACE, which is designed to identify disconnected regions
of a biochemical parameter space and characterise them by estimating
their dimensions. HYPERSPACE does not perform parameter inference and is
therefore not directly comparable with algorithms like adaptive Hamiltonian
Monte Carlo as used in Maud, but could plausibly be adapted into such an
algorithm, for example by supplementation with importance sampling (see
@vehtariParetoSmoothedImportance2022). However, any such algorithm would
be limited by HYPERSPACE's dependency on a random walk Metropolis Hastings
algorithm. This kind of algorithm's performance is known to scale poorly
compared with Hamiltonian Monte Carlo algorithms as the number of parameters
increases: see @mangoubiDoesHamiltonianMonte2018. Since Maud aims to fit models
with hundreds of parameters, this means that it is unlikely that HYPERSPACE can
directly be used to improve Maud's efficiency.

>>>>

<!--  :::{.callout-important title="Changes"} -->
While we have not yet observed this in practice, we expect that Maud is capable
of accurately sampling mildly multi-modal posterior distributions, i.e. those
for which regions of parameter space with high probability density are not
very sharply separated. This is because this ability depends primarily on the
underlying inference algorithm, and adaptive Hamiltonian Monte Carlo is known
to be able to sample many such posterior distributions. See, for example,
@mangiolaSccompRobustDifferential2023, which reports the successful use of
adaptive Hamiltonian Monte Carlo to sample posterior distributions involving
mixtures of Gaussian distributions, which are typically multi-modal.

@@

@zamora-silleroEfficientCharacterizationHighdimensional2011 introduces
the method HYPERSPACE, which is designed to identify disconnected regions
of a biochemical parameter space and characterise them by estimating
their dimensions. HYPERSPACE does not perform parameter inference and is
therefore not directly comparable with algorithms like adaptive Hamiltonian
Monte Carlo as used in Maud, but could plausibly be adapted into such an
algorithm, for example by supplementation with importance sampling (see
@vehtariParetoSmoothedImportance2022). However, any such algorithm would
be limited by HYPERSPACE's dependency on a random walk Metropolis Hastings
algorithm. This kind of algorithm's performance is known to scale poorly
compared with Hamiltonian Monte Carlo algorithms as the number of parameters
increases: see @mangoubiDoesHamiltonianMonte2018. Since Maud aims to fit models
with hundreds of parameters, this means that it is unlikely that HYPERSPACE can
directly be used to improve Maud's efficiency.
<!-- ::: -->
